---
title: "Assignment 6, task 3"
author: "Quin Smith"
date: "12/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(effsize)
library(broom)
```

```{r}
#read in data
lizards <- read_csv(here("data", "lizards.csv"))

#lizard subset that includes female CNUNs and CNTEs
lizards_subset <- lizards %>% 
  select(spp, sex, total_length) %>% 
  filter(spp %in% c("CNUN", "CNTE"),
         sex == "F") %>% 
  mutate(common_name = case_when(
    spp == "CNUN" ~ "Desert Grassland lizard",
    spp == "CNTE" ~ "Colorado Checkered lizard"
  ))
```

### Visualization

```{r, fig.cap= ""}
#create visualization
lizards_subset %>% 
  ggplot(aes(x = total_length)) +
  geom_histogram(bins = 10) +
  labs(x = "Total Length (mm)", y = "Count") +
  facet_wrap(~common_name, scales = "free") +
  theme_bw()
```

```{r}
#create summary table for mean and sd of total_length across species
sumry_table <- lizards_subset %>% 
  group_by(spp) %>% 
  summarize(mean = mean(total_length, na.rm = TRUE),
            sd = sd(total_length, na.rm = TRUE))

#create vectors
cnun_vec <- lizards_subset %>%
  filter(spp == "CNUN") %>%
  pull(total_length)

cnun_vec <- cnun_vec[!is.na(cnun_vec)]

cnte_vec <- lizards_subset %>% 
  filter(spp == "CNTE") %>% 
  pull(total_length)

cnte_vec <- cnte_vec[!is.na(cnte_vec)]

#exploratory plots
qqnorm(cnun_vec)
qqnorm(cnte_vec)

hist(cnun_vec)
hist(cnte_vec)

#Two sample t-test
liz_tt <- t.test(cnun_vec, cnte_vec)

liz_tt_tidy <- tidy(liz_tt)

#Cohen-D
liz_cd <- cohen.d(cnun_vec, cnte_vec)
```
On average, female desert grassland lizards are longer than female Colorado checkered lizards (`r round(sumry_table[2,2], 2)` $\pm$ `r round(sumry_table[2,3], 2)` and `r round(sumry_table[1,2], 2)` $\pm$ `r round(sumry_table[2,3], 2)` cm, respectively; mean $\pm$ 1 standard deviation). While the absolute difference in means is `r round(sumry_table[2,2], 2) - r round(sumry_table[1,2], 2)` cm (a `r round(sumry_table[2,2], 2) - r round(sumry_table[1,2], 2)/((r round(sumry_table[2,2], 2) + r round(sumry_table[1,2], 2))/2)*100, 2)`% difference), the difference in means is significant (Welch's two-sample t-test: t(`r round(liz_tt_tidy$parameter, 2)`) = `r round(liz_tt_tidy$statistic, 2)`, p < 0.001), and the effect size is moderate (Cohen's *d* = `r round(liz_cd$estimate, 2)`).







